# (APPENDIX) Appendix {-}

# Connecting to NAPMD

NAPMD can be accessed via:

- an API (requiring an API username and password), or 
- Postgres connection (requiring a database username and password).

The instructions that follow are aimed primarily at `R` users. To use `R`, you will need to install both [R](https://www.r-project.org/) and [RStudio](https://posit.co/download/rstudio-desktop/).

## Using API with R `napmdtools` package

`R` functions that interface with the NAPMD API can be found in the `napmdtools` package. See the package repository and instructions [here][napmdtools-repo].

## Via DB connection with R `RPostgreSQL` package

Connect to the database with the following:

```{r example-r-dbconn, eval = F, echo = T}
# Load RPostgreSQL library
require("RPostgreSQL")

# Load the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

## create your password
## Either (NOTE: this window can open in the background, so you might need to minimise RStudio to see it.
postgres_pwd <- askpass::askpass(prompt = "Please enter your password: ")
## OR
# postgres_pwd <- "<your_password>"

## Create a connection to the postgres database 
ch <- dbConnect(drv, dbname = "postgis_car",
                host = "<server_name>", port = 5432,
                user = "<user_name>", password = postgres_pwd)

remove(postgres_pwd)
```

Once connect, use the `dbGetQuery` function to query with the database using SQL. For example:

```{r example-r-dbquery, eval = F, echo = T}

df.mons  <- dbGetQuery(ch, "SELECT * from air_pollution_monitors.ap_monitor_data_master 
                       where state = 'SA'
                       limit 100;")

```

For geospatial work, you can read the locations tables directly into an `sf` or `SpatVector` (`terra`) object through the `rpostgis` package:

```{r example-rpostgis, eval = F, echo = T}
library(rpostgis)
# read as sf object
sf.mons <- rpostgis::pgGetGeom(ch, 
                               c("air_pollution_monitors", "ap_monitor_locations_act"), 
                               geom = "geom_gda94",
                               returnclass = "sf")
# or as terra SpatVector
v.mons <- rpostgis::pgGetGeom(ch, 
                              c("air_pollution_monitors", "ap_monitor_locations_act"), 
                              geom = "geom_gda94",
                              returnclass = "terra")

```


## Via pgAdmin 4

1. Download and install PGAdmin 4 <https://www.pgadmin.org/download/>
2. Run PGAdmin 4
3. In the left sidebar, right-click on `Servers` -> `Create` -> `Server...`
4. Name the Server in the General Tab
5. In the Connections tab fill in the following details, replacing <Host name/address>, <Username> and <Password> with the credentials given to you. Then save the changes.

You may then navigate using the tree in the left sidebar. Find the `air_pollution_monitors` schema in the Schemas list.

To run SQL queries, open the Query Tool (`Tools` > `Query Tool` in the top menu, or right-click a schema or table and click `Query Tool`). Run your query with the `Execute/Refresh` button (shortcut key `F5`) above the top-right pane. Save query results with the `Save results to file` button (shortcut key `F8`).


